---
title: Curcumin Treated T-C1 Cell Proliferation (MTS)
author: Davit Sargsyan
output:
  html_document:
    theme: null
    highlight: null
    media: screen
    css: style.css
    toc: true
---
Date: `r Sys.time()`

# Design

T-C1 cells were dosed with serially diluted (.../2) Curcumin. Each of the 3 96-well plate contained 6 replicas for each of the 9 doses, plus DMSO control and medium control (background). Plates were analyzed at 24, 48 and 72 hours. Each plate was red twice. The results are presented below.

# Analysis

Data was analyzed using R 3.2.1. The estimates and plots were obtained using my robust 4-parameter curve fitting R package 'ricf'. 

## Raw Data

```{r raw, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE}
options(scipen = 999,
        warn = -1)

setwd("C:\\Users\\dsargsy\\Desktop\\PhD\\Lab")
require(data.table)
require(reshape2)
require(knitr)
require(ricf)

dilut.nM <- rev(c(format(100000/2^(0:8),
                         scientific = TRUE, 
                         digits = 3),
                  "DMSO Only"))

plates <- fread("TC1_Curcumin_MTS_Plates_Exp1.csv", sep = ",", header = TRUE)
maps <- fread("TC1_Curcumin_MTS_Map_Exp1.csv", sep = ",", header = TRUE)

d1 <- reshape(data = plates, 
              varying = list(names(plates)[5:ncol(plates)]),
              direction = "long",
              v.names = "Value")

d1$Treatment <- reshape(data = maps, 
                        varying = list(names(maps)[4:ncol(maps)]),
                        direction = "long",
                        v.names = "Treatment")$Treatment

d1$time <- NULL

# Median of background, i.e. medium only, per plate
d1[, bgrnd := median(Value[Treatment == "None"]),
   by = list(Plate, Reading)]

# Subtract background
d1[, no.bgrnd := Value - bgrnd]

# Separate treated cells
trt <- subset(d1, !(Treatment %in% c("None", "MTS")))
plot(trt$Value ~ rev(trt$Treatment), 
     col = as.numeric(factor(trt$Hour)),
     xlab = "Concentration",
     ylab = "Value",
     main = "All Plates, Raw Data (24, 48 and 72 Hour)",
     xaxt = "n")

dd <- dcast(trt, Row + Treatment ~ Reading + Hour, value.var = "Value")
dd$Treatment[dd$Treatment == "DMSO"] <- 9

# Analysis
dd$dilut <- 100000/2^as.numeric(as.character(dd$Treatment))
m1 <- f.robicf(data.frame(conc = log10(dd$dilut),
                          dd[, c(3, 6, 4, 7, 5, 8)]))
s1 <- summary(m1, export = TRUE)

# for (i in 1:length(s1)) {
#   cat(kable(s1[[i]],
#           format="html",
#           row.names = TRUE,
#           digits = 3,
#           align = "l",
#           caption = names(s1)[i]))
# cat("<br><br>")
# }
# 
# plot(m1,
#      xlab = "Curcumin Concentration (nM)",
#      ylab = "Raw Cell Counts (Units?)",
#      smooth  = TRUE,
#      xmarks = dilut.nM)

lines(m1,
      smooth = TRUE,
      xlab = "Log of Curcumin Concentration (nM)",
      ylab = "Raw Cell Counts (Units?)",
      main = "All Plates, Raw Data (24, 48 and 72 Hour)")
```

## Normalized Data

Norm. Response = 100*(Treatment - Median(Plate Background))/(Row DMSO - Median(Plate Background))

```{r norm, echo=FALSE, results='asis', message=FALSE, error=FALSE, warning=FALSE}
options(scipen = 999,
        warn = -1)

trt[, DMSO := no.bgrnd[Treatment == "DMSO"],
    by = list(Plate, Reading, Row)]
trt[, Normilized := 100*no.bgrnd/DMSO]

trt <- subset(trt, Treatment != "DMSO")

plot(trt$Normilized ~ rev(trt$Treatment), 
     col = as.numeric(factor(trt$Hour)),
     xlab = "Concentration",
     ylab = "Value",
     main = "All Plates, Normalized (24, 48 and 72 Hour)",
     xaxt = "n")

dd <- dcast(trt, Row + Treatment ~ Reading + Hour, value.var = "Normilized")

# Analysis
dd$dilut <- 100000/2^as.numeric(as.character(dd$Treatment))
m1 <- f.robicf(data.frame(conc = log10(dd$dilut),
                          dd[, c(3, 6, 4, 7, 5, 8)]),
               ecf = c(20, 50, 80))
s1 <- summary(m1, export = TRUE)

for (i in 1:length(s1)) {
  cat(kable(s1[[i]],
          format="html",
          row.names = TRUE,
          digits = 3,
          align = "l",
          caption = names(s1)[i]))
cat("<br><br>")
}

plot(m1,
     xlab = "Curcumin Concentration (nM)",
     ylab = "Normalized Response (% DMSO)",
     smooth  = TRUE,
     xmarks = dilut.nM[-1])

lines(m1,
      smooth = TRUE,
      xlab = "Log of Curcumin Concentration (nM)",
      ylab = "Normalized Response (% DMSO)")
```

# Discussion

1. Although data followed 4-parameter model well for 24 and 72 hour read-outs, 48 hour read-out had major issues. 

2. Variance decreased from higher to lower doses but the robust model assumes homogeneity of variance across concentrations. Although this might not have influenced the estimates in this case, generally a transformation of the response might be needed to stabilize variance.

3. Hook effect was observed at lower doses. A model that can describe the hook might be more appropriate, especially if the hook effect becomes more pronounced.

4. Smaller intervals between the doses can be beneficial as we can get smoother curves around IC50 values and decrease standard errors of the estimates.